---
title: "Analysis of the ToothGrowth data in R"
author: "Ray Qiu"
date: "October 22, 2015"
output: 
    pdf_document:
        latex_engine: xelatex
subtitle: Statistical Inference Course project part 2
---

## Overview
In the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package, and try to draw some conclusions based on statistical inference.

---

#### Load required R libraries
```{r message=FALSE}
library(ggplot2)
library(dplyr)
```

## Load the ToothGrowth data and perform some basic exploratory data analysis. Provide a basic summary of the data.
```{r}
data("ToothGrowth")
str(ToothGrowth)
head(ToothGrowth)
tail(ToothGrowth)
summary(ToothGrowth)
```
#### Convert dose to factors
```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```
#### Some exploratory plots
```{r}
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
    ggtitle("Teeth Length per Dosage Level") +
    geom_boxplot(aes(fill = dose))
ggplot(ToothGrowth, aes(x = supp, y = len)) + 
    ggtitle("Teeth Length per Supplement Type") +
    geom_boxplot(aes(fill = supp))
```

## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
#### Test whether supplement types have effects on teeth length
```{r}
t.test(len ~ supp, data = ToothGrowth)
```
The test result shows that p-value is 0.06063, which is greater than 0.05; And 95 percent confidence interval values contain 0. Based on this we cannot reject the null hypothesis that different supplement types have no effect on teeth length.

#### Separate data into three dose groups where each group contains a pair of doses, then test each dose pair to see whether different doses have different effects on teeth length.
```{r}
doseGroup1 <- ToothGrowth %>% filter(dose != 2.0)
doseGroup2 <- ToothGrowth %>% filter(dose != 1.0)
doseGroup3 <- ToothGrowth %>% filter(dose != 0.5)
```

#### Test doseGroup1
```{r}
t.test(len ~ dose, data = doseGroup1)
```

#### Test doseGroup2
```{r}
t.test(len ~ dose, data = doseGroup2)
```

#### Test doseGroup3
```{r}
t.test(len ~ dose, data = doseGroup3)
```
The test result shows that p-value is close to 0; And 95 percent confidence interval values don't contain 0. Based on this we can reject the null hypothesis, and conclude that increasing dosage will increase teeth growth.

## State your conclusions and the assumptions needed for your conclusions. 

#### Conclusions

* Supplement type has no effect on teeth growth.

* Increasing dosage level leads to increased teeth growth.

#### Assumptions

* The experiment design suggest that each Guinea pig was randomly assigned to a combination of dosage and supplement type, so the tests use the independent samples methodology.

* The sample of 60 Guinea pigs is representative of all Guinea pigs, so we can generalize the test conclusion to the population.

* For t-test regarding tooth length per supplement type, the variances are assumed to be different for the two groups being compared.  For t-tests regarding teeth length per dosage level, we assume a constant variance across the groups.
