---
title: "Statistical Inference (Quiz 2)"
author: "Ray Qiu"
date: "October 14, 2015"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    keep_md: yes
---

1. What is the variance of the distribution of the average an IID draw of n observations from a population with mean μ and variance σ2.

    * σ ^ 2 / n

2. Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?
```{r}
μ <- 80
σ <- 10
round(pnorm(70, μ, σ), 2) * 100
```

3. Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?
```{r}
μ <- 1100
σ <- 75
round(qnorm(0.95, μ, σ), 0)
```

4. Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?
```{r}
μ <- 1100
σ <- 75
n <- 100
S <- sqrt(σ ^ 2 / n)
round(qnorm(0.95, μ, S), 0)
```

5. You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?
```{r}
p <- 0.5
n <- 5
μ <- 0.5
round(pbinom(3, 5, p, lower.tail = FALSE), 2) * 100
```

6. The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?
```{r}
n <- 100
σ <- 10
μ <- 15
S <- σ / sqrt(n)
round((pnorm(16, μ, S) - pnorm(14, μ, S)), 2) * 100
```

7. Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?
* With a large enough sample like 1000 observations, the sample mean should be the population mean, which is 0.5.

8. The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?
```{r}
q <- 10
lambda = 5 * 3
round(ppois(q, lambda), 2)
```


